Cluster Difference Imaging Photometric Survey. III. An Open Cluster Spanning 500\,Parsecs

### 2021-04-01 cluster_paper/ms.tex
update ms

% gaia target sample numbers
\newcommand{\nkinematic}{3{,}298\ } % "fullfaint" kinematic sample, core+halo (CG18+KC19+M21). from any call of earhat.helpers._get_fullfaint_dataframes
\newcommand{\nnbhd}{13{,}843\ } % make_compstar_NGC_2516_sourcelist.py
\newcommand{\ncore}{1{,}106}  % "fullfaint" kinematic sample, CG18. from any call of earhat.helpers._get_fullfaint_dataframes
\newcommand{\nhalo}{2{,}2192} % "fullfaint" kinematic sample, KC19+M21. from any call of earhat.helpers._get_fullfaint_dataframes

% from drivers.cluster_rotation.get_auto_rotation_periods.py;  checked with plot_target_vs_actual_list_sanity_checks.py
\newcommand{\ncdips}{2{,}205\ }
\newcommand{\nkinematicrpltsixteen}{2{,}270\ }
\newcommand{\nkinematicCGandKC}{2{,}236\ }

% cdips results numbers
\newcommand{\numcdipslcs}{671{,}901\ } % drivers.yield_calculation.count_cdips_lcs.py; 2021.01.13
\newcommand{\numsouthernuniqlcs}{483{,}407\ } % drivers.yield_calculation.count_cdips_lcs.py; 2021.01.13

\newcommand{\ncalibration}{9{,}619{,}784\ } % /nfs/phtess2/ar0/TESS/PROJ/lbouma/CDIPS_LCS/counts/check_sum, 2021.02.02
\newcommand{\nnbhdcaliblc}{9{,}790\ } % get_auto_rotation_periods.py
\newcommand{\nnbhdcalibstar}{1{,}987\ } % get_auto_rotation_periods.py

\newcommand{\nautorotdenominator}{1{,}641\ } % plot_autorotation.py, w/ runid=NGC_2516. For nequal<=0, nclose<=1, LSP>0.08. See /doc/20210205_nequal_and_Prot_vs_color_cleanup.txt
\newcommand{\nautorotnumerator}{987\ } % plot_autorotation.py, For nequal<=0, nclose<=1, LSP>0.08. See /doc/20210205_nequal_and_Prot_vs_color_cleanup.txt
\newcommand{\nautorotnumeratormatching}{507\ } % plot_autorotation.py, LSP=SPDM period
\newcommand{\ncompstardenominator}{820\ } % plot_autorotation.py, w/ runid=compstar_NGC_2516. For nequal<=0, nclose<=1, LSP>0.08. 
\newcommand{\ncompstarnumerator}{365\ } % plot_autorotation.py
\newcommand{\ncompfrac}{44.5\%} % plot_autorotation.py
\newcommand{\nautovscompstardenominator}{705\ } % plot_autorotation.py, w/ runid=compstar_NGC_2516. For nequal<=0, nclose<=1, LSP>0.08. 
\newcommand{\nautovscompstarnumerator}{374\ } % plot_autorotation.py
\newcommand{\nautofrac}{53.1\%} % plot_autorotation.py

@@

We have recently been making TESS light curves of stars reported to be
members of coeval populations (CDIPS, \citealt{bouma_cdipsI_2019}).
Our analysis of TESS Sectors 1-13 yielded light curves of
\numsouthernuniqlcs candidate cluster members in the Southern Ecliptic
hemisphere, which are available on
MAST\footnote{\url{https://archive.stsci.edu/hlsp/cdips}}.

@@

\begin{figure*}[t]
	\begin{center}
		\leavevmode
			%\includegraphics[width=0.85\textwidth]{f2a.pdf}
		\subfloat{
			\includegraphics[width=0.49\textwidth]{f2a.pdf}
			\includegraphics[width=0.49\textwidth]{f2b.pdf}
		}
	\end{center}
	\vspace{-0.7cm}
  \caption{ {\bf HR diagrams of NGC\,2516.}
    {\it Left}: Photometric data from Gaia EDR3; no reddening
    correction has been applied.  The core (black) shows a
    main-sequence and turnoff consistent with stars of a fixed age and
    metallicity.  The halo (blue) is similar, though the binary
    sequence is less defined.  The faintest M dwarfs in the core and
    halo are brighter than field stars (gray) with the same color,
    consistent with the cluster M dwarfs not having yet reached the
    main-sequence.  {\it Right}: PARSEC isochrones ([Fe/H]=+0.06) are
    overplotted at intervals of $\log_{10}(t/\mathrm{yr})=\{8, 8.25,
    8.5\}$.  The arrow represents the average reddening correction
    applied to the models.  The opacity of the M-dwarf model
    atmospheres is less well-calibrated for $M_\star \lesssim
    0.45M_\odot$ ($\approx$M2V; $Bp-Rp\approx$2.2); these model values
    are shown with dashed lines.
    \label{fig:hr}
  }
\end{figure*}

@@

        \vspace{-0.5cm}
		\subfloat{
			\includegraphics[width=0.49\textwidth]{f3c.pdf}
			\includegraphics[width=0.49\textwidth]{f3d.pdf}
		}
	\end{center}
	\vspace{-0.7cm}
  \caption{ {\bf TESS rotations periods and dereddened Gaia colors for \cn.}
    {\it Top row}: \cn\ split into its core and halo.  Subset
    $\mathcal{A}$ ({\it left columns}) and $\mathcal{B}$ ({\it right
    columns}) undergo successive degrees of automated cleaning,
    described in Section~\ref{subsubsec:cluster}.  The Pleiades
    \citep[120\,Myr;][]{rebull_rotation_2016a} and Praesepe
    \citep[650\,Myr;][]{douglas_poking_2017} are shown for reference.
    Stars in the core and the halo of NGC\,2516 appear
    gyrochronologically coeval with the Pleiades at $(Bp-Rp)_0$$<$1.2.
    There is a narrow range of color, from $1.2<(Bp-Rp)_0<1.7$, in
    which some NGC\,2516 members appear slightly elevated with respect
    to the Pleiades, suggesting an older age.
    {\it Bottom row}: \cn\ colored by binarity indicators.  Sources
    with $\mathrm{RUWE}>1.2$ appear in red (astrometric binaries; 11\%
    of stars).
    Overplotted in orange are sources at least 0.3\,mag above an
    empirically fit cluster isochrone (photometric binaries; 20\% of stars).
		\label{fig:rot}
	}
\end{figure*}

We began our rotation analysis by considering all \nkinematic\
candidate members of \cn.  For each source, we first retrieved all
available CDIPS light curves, on a per-sector basis.  This yielded
\ncdips stars with at least one sector from a CDIPS light curve, all
brighter than $Rp=16$\,mag.  \nkinematicrpltsixteen\ of the stars from
the source list have $Rp<16$.  The difference is comprised of 34 stars
unique to \citet{meingast_2021} which were not available at the time
of the CDIPS reductions and stars falling on the TESS chip gaps.  The
$Rp=16$\,mag cutoff imposed during the CDIPS processing corresponds
roughly to $(Bp-Rp)_0$ of 2.2, or a spectral type of $\approx$M2V, at
the distance of NGC\,2516.

We removed systematic trends shared between all light curves on each
CCD in each sector individually, and stitched together the resulting
light curves before searching for the periodicity.  Details regarding
our detrending approach are discussed in
Appendix~\ref{app:detrending}.  After applying the initial detrending
step aimed at removing systematic trends, we proceeding with a few
small cleaning steps aimed at improving the purity of the rotation
period measurements: we masked 0.7 days at the beginning and end of
each spacecraft orbit, and ran a sliding standard-deviation rejection
window over the light curve, which removed any outlying points within
$\pm3\times$MAD of the median in each window.

We then measured the rotation period from the resulting light curve.
We used the CDIPS aperture radius that, based on theoretical
expectations, was expected to give the optimal balance between light
from the target and background-light \citep{Sullivan_2015}.  This
typically resulted in an aperture radius of either 1 or 1.5 pixels.
To measure the periods, we used the periodogram implementations in
\texttt{astrobase}, in particular the Stellingwerf phase-dispersion
minimization periodogram, along with the more traditional Lomb-Scargle
\citep{lomb_1976,stellingwerf_period_1978,scargle_studies_1982,stellingwerf_period_2011,bhatti_astrobase_2018}.
We recorded the top five periodogram peaks from each method, and their
corresponding powers.  Finally, as a check on crowding, we also
recorded the number of stars within the aperture of equal brightness
to the target stars, and of brightness with 1.25 and 2.5 TESS
magnitudes of the target star.

The top row of Figure~\ref{fig:rot} shows the resulting rotation
periods, following three different sets of cleaning criteria.  The
first and second sets of criteria were entirely automated, and yielded
subsets we denote $\mathcal{A}$ and $\mathcal{B}$.  For both subsets,
we considered light curves for which the peak Lomb-Scargle periodogram
period was below 15 days, the normalized Lomb-Scargle power exceeded
0.08, and for which no equal-brightness or greater companions were
within the aperture.  Beyond requiring that the target star be the
brightest star in the aperture, we also required that at most one
companion with brightness exceeding one tenth of the target star's
brightness be present in the aperture according to the Gaia DR2 source
catalog.  For set $\mathcal{A}$, this yielded \nautorotnumerator light
curves, out of \nautorotdenominator stars for which rotation periods
might plausibly have been detected.  For set $\mathcal{B}$, we 
additionally required that the Lomb-Scargle and Stellingwerf
phase-dispersion methods yielded identical periods to within 10\%.
This yielded \nautorotnumeratormatching light curves. 
While subset $\mathcal{A}$ is more complete than subset $\mathcal{B}$,
this completeness comes at the expense of greater contamination.
% For instance, stars above the slow sequence (possible field
% contaminants) for the most part disappear in subset $\mathcal{B}$

The comparison to the Pleiades and Praesepe is made using the rotation
periods and dereddened colors from Table~5 of
\citet{curtis_rup147_2020}, which used data from
\citet{rebull_rotation_2016a} and \citet{douglas_k2_2019}
respectively.
A few conclusions can be drawn by comparing the \cn\ rotation periods
with those of the Pleiades and Praesepe.  The first is that in both
stages of cleaning, the core and halo of \cn\ each show
gyrochronological ``slow sequences''.  These sequences appear to be
coeval with each other, and at first order appear to overlap
with the Pleiades sequence.

% Second, subset $\mathcal{A}$ is more complete than subset
% $\mathcal{B}$.  This completeness comes at the expense of greater
% contamination.  The slow sequence has a factor of $\sim$twice as many
% stars in subset $\mathcal{A}$ than $\mathcal{B}$.  However stars above
% the slow sequence, which could be field contaminants, for the
% most part disappear in subset $\mathcal{B}$.  The ``void'' underneath
% the slow sequence is also more pronounced in the cleaner subset,
% though it seems to not be as empty as in the Pleiades.

%FIXME this belongs in the DISCUSSION!!
%FIXME this belongs in the DISCUSSION!!
%FIXME this belongs in the DISCUSSION!!
%FIXME this belongs in the DISCUSSION!!
Comparing the slow sequence of the Pleiades and \cn\ 
more closely, the top row of Figure~\ref{fig:rot} shows that the two
sequences
overlap from roughly $0.5<(Bp-Rp)_0<1.2$.  At redder colors
($1.2<(Bp-Rp)_0<1.7$) the dispersion in rotation periods increases.
The maximum rotation periods seen in \cn, in both subsets
$\mathcal{A}$ and $\mathcal{B}$ extend up to $\approx$11 days, rather
than the $\approx$8.5day limit seen in the Pleiades.  This is
consistent with \cn\ being slightly older than the Pleiades.  Fitting
a model to substantiate this statement \citep[{\it e.g.}, any
of][]{mamajek_improved_2008,angus_toward_2019,spada_competing_2020} is
an exercise in tautology, since the aforementioned models are empirically fit using
the Pleiades and Praesepe.  In other words, at $(Bp-Rp)_0=1.35$, we
see rotation periods in \cn\ ranging from 7 to 10 days.  The 7 day
periods overlap with the Pleiades; the 10 day periods do not.  Fitting
formulae from the previously cited studies give ages for
a star with $(Bp-Rp)_0=1.35, (B-V)_0=1.10, P=7\,{\rm day}$ of 204 Myr,
316 Myr, and 107 Myr respectively.  Therefore, only the
\citet{spada_competing_2020} model is well-calibrated at these ages
and colors.  For an 8 day rotation period of a star with the same
color, this model quotes a 193 Myr age; at 9 days, 388 Myr.
Given the model uncertainty, we prefer the relative statements that
\cn\ appears slightly gyrochronologically older than the Pleaides, and
much younger than Praesepe.

The bottom row of Figure~\ref{fig:rot} shows the same data, with the
points colored by indicators of binarity.  To assess astrometric
binarity, we used the renormalized unit weight error (RUWE) from Gaia
EDR3.  For plotting purposes, we labeled anything with RUWE exceeding
1.2 as an astrometric binary (11\% of the overall cluster sample).  A
cautionary point is that poor astrometric fits can also be caused by
photometric variability.  To assess photometric binarity, we fitted a
spline to Figure~\ref{fig:hr} by hand, and labelled any points
brighter than 0.3\,mag above the cluster sequence as photometric
binaries.  This included 20\% of the overall cluster sample.  Many of
the resulting candidate binaries, though not all of them, appear
underneath the ``slow sequence''.  We speculate on possible
explanations in Section~\ref{disc:rotn}.

@@

\begin{figure*}[t]
	\begin{center}
		\subfloat{
			\includegraphics[width=0.5\textwidth]{f8a.pdf}
			\includegraphics[width=0.5\textwidth]{f8b.pdf}
		}
	%		\includegraphics[width=0.7\textwidth]{f8a.pdf}
	\end{center}
	\vspace{-0.7cm}
  \caption{ 
  {\bf Members of \cn\ exist hundreds of parsecs from the core,
  and up to $\approx$5\kms\ from the core in tangential velocity.}
  Halo members that meet both kinematic and rotational membership
  criteria (Subset~$\mathcal{A}$) are shown in blue.  Halo members for which rotation periods
  should have been detectable, but were not detected, are shown in
  orange.  
  The plotted stars are restricted to $0.5<(Bp-Rp)_0<1.2$.
  The {\it left panels} show galactic coordinates and
  2D-tangential velocity space, after
  correcting for projection effects
  (Appendix~\ref{app:vproj}).
  The histogram on the right bins versus 3D separation from the core
  and 2D velocity difference.  Bins have width 25\,pc in position, and
  1.0\,km\,s$^{-1}$ in velocity.  {\bf Uncertainties in the fraction
  are taken as the quadrature sum of the rel unc of the numerator and
  denominator (FIXME)}
  \label{fig:physical_x_rotn}
	}
\end{figure*}

@@

How far away from the core, in position and velocity space, does the
halo extend?  We can explore this by crossmatching the stars with
detected rotation periods against our original target list of
\nkinematic Gaia DR2 kinematic candidate members.  For a fair
comparison, we consider only the stars with $0.5<(Bp-Rp)_0<1.2$ for
which our TESS pipeline successfully made detrended light curves.  In
other words---the base sample is the stars for which we could have
plausibly detected a rotation period.  Faint stars with $Rp>16$ were
beyond our light curve selection limit, and crowded stars ({\it e.g.,}
near saturated stars in the cluster center) are similarly omitted.  We
selected the color limit above based on Figure~\ref{fig:rot}, since
the density of period detections seems to decrease once $(Bp-Rp)_0
\gtrsim 1.25$, {\it i.e.}, for spectral types later than $\approx$K4V.
Our expectation is that completion effects become important for
fainter stars.

@@

Gaia-ESO selected candidate \cn\ members to observe with the GIRAFFE
and UVES spectrographs based on previously reported literature members
and publicly available photometry.  Since the existence of the \cn\
halo was not known at the time of target selection, very few ``halo''
stars are in the sample.  For \cn, \citet{randich_gaiaeso_2018}
ultimately reported stellar parameters (including lithium equivalent
widths and metallicities) for 796 stars that they considered possible
\cn\ members.  Cross-matching against our kinematic list of
\nkinematic candidate members by position and imposing a 0$\farcs$5
maximum separation limit yields 492 kinematic members for which a
Gaia-ESO spectrum was acquired and stellar parameters were reported.
15 of these (all with $(Bp-Rp)_0 > 2.0$) are spurious matches based on
the Gaia color and GES effective temperature, and we remove them.
This yields 477 stars, of which 436 are in the core, and 41 are in the
halo.  The lop-sided ratio is due to the Gaia-ESO selection function.
We verified by directly querying the ESO archive that
\citet{randich_gaiaeso_2018} indeed included $>99\%$ of the available
Gaia-ESO spectra as part of their \cn\ sample, and so we opted to use
their results rather than repeating their analysis.

The GALAH DR3 target selection is discussed in depth by
\citet{buder_galah_2020}.  The relevant aspects for our analysis are
that the ``main'' survey targeted $12<V<14$ stars at
$\delta<+10^\circ$, provided the stars were at least ten degrees from
the galactic plane.  The survey is spatially
inhomogeneous\footnote{See the footprint at
\url{https://www.galah-survey.org/news/announcing-galah-dr3}}. Special
targeting for stars in the TESS southern continuous viewing zone, and
for known open cluster members was also performed.  We identified the
candidate \cn\ members for which spectra had been obtained by
searching the \texttt{GALAH\_DR3\_main\_allstar\_v1} catalog, after
excluding stars with the stellar parameter bit flags 1, 2, 3.  This
excludes spectra with unreliable broadening, low S/N, and unreliable
wavelength solutions (see Table~6 of \citealt{buder_galah_2020}).
Since our main focus is measuring equivalent widths for the
Li\,6708\,\AA doublet, these were the most relevant bit flags.  Of our
\nkinematic candidate \cn\ members, 107 had spectra in GALAH DR3.  51
were in the ``core''; 56 were in the ``halo''.  78 had ``finite''
lithium, i.e., no lithium flag set.

@@

\subsection{Rotation in NGC\,2516}
\label{disc:rotn}

Gotta cite Fritzewski+2020.  (Maybe even do the Blanco1, Pleiades,
NGC2516, PscEri concatenation?)

  \paragraph{Slow sequence and fast sequence}
  %Why is there a slow and fast sequence?
  Slow: Skumanich spindown, Weber and Davis magnetic braking.
  This doesn't explain everything.
  Originally described as ``interface / convective'' dynamo for each.
  Now disfavored (but why?).
  Core envelope decoupling handwaves away the fast rotators? (Irwin,
  Aigrain, Monitor).
  Another fast rotator explanation: Naoz's merging of giant planets
  (could explain some fraction of the fast rotators -- not all of them).
  Tim Brown, ``metastable dynamo''?

  Perhaps most compelling: some kind of binary related explanation.
  Gillen+20 in Blanco-1 found most of them to be photometric binaries.
  Simonian+19 in the Kepler field found most rapid rotators to have
  photometric excesses.

>>>>

% gaia target sample numbers
\newcommand{\nkinematic}{3{,}298} % "fullfaint" kinematic sample, core+halo (CG18+KC19+M21). from any call of earhat.helpers._get_fullfaint_dataframes
\newcommand{\nnbhd}{13{,}843} % make_compstar_NGC_2516_sourcelist.py
\newcommand{\ncore}{1{,}106}  % "fullfaint" kinematic sample, CG18. from any call of earhat.helpers._get_fullfaint_dataframes
\newcommand{\nhalo}{2{,}2192} % "fullfaint" kinematic sample, KC19+M21. from any call of earhat.helpers._get_fullfaint_dataframes

% from drivers.cluster_rotation.get_auto_rotation_periods.py;  checked with plot_target_vs_actual_list_sanity_checks.py
\newcommand{\ncdips}{2{,}205}
\newcommand{\nkinematicrpltsixteen}{2{,}270}
\newcommand{\nkinematicCGandKC}{2{,}236}

% cdips results numbers
\newcommand{\numcdipslcs}{671{,}901} % drivers.yield_calculation.count_cdips_lcs.py; 2021.01.13
\newcommand{\numsouthernuniqlcs}{483{,}407} % drivers.yield_calculation.count_cdips_lcs.py; 2021.01.13

\newcommand{\ncalibration}{9{,}619{,}784} % /nfs/phtess2/ar0/TESS/PROJ/lbouma/CDIPS_LCS/counts/check_sum, 2021.02.02
\newcommand{\nnbhdcaliblc}{9{,}790} % get_auto_rotation_periods.py
\newcommand{\nnbhdcalibstar}{1{,}987} % get_auto_rotation_periods.py

\newcommand{\nautorotdenominator}{1{,}641} % plot_autorotation.py, w/ runid=NGC_2516. For nequal<=0, nclose<=1, LSP>0.08. See /doc/20210205_nequal_and_Prot_vs_color_cleanup.txt
\newcommand{\nautorotnumerator}{987} % plot_autorotation.py, For nequal<=0, nclose<=1, LSP>0.08. See /doc/20210205_nequal_and_Prot_vs_color_cleanup.txt
\newcommand{\nautorotnumeratormatching}{507} % plot_autorotation.py, LSP=SPDM period
\newcommand{\ncompstardenominator}{820} % plot_autorotation.py, w/ runid=compstar_NGC_2516. For nequal<=0, nclose<=1, LSP>0.08. 
\newcommand{\ncompstarnumerator}{365} % plot_autorotation.py
\newcommand{\ncompfrac}{44.5\%} % plot_autorotation.py
\newcommand{\nautovscompstardenominator}{705} % plot_autorotation.py, w/ runid=compstar_NGC_2516. For nequal<=0, nclose<=1, LSP>0.08. 
\newcommand{\nautovscompstarnumerator}{374} % plot_autorotation.py
\newcommand{\nautofrac}{53.1\%} % plot_autorotation.py

@@

We have recently been making TESS light curves of stars reported to be
members of coeval populations (CDIPS, \citealt{bouma_cdipsI_2019}).
Our analysis of TESS Sectors 1-13 yielded light curves of
\numsouthernuniqlcs\ candidate cluster members in the Southern Ecliptic
hemisphere, which are available on
MAST\footnote{\url{https://archive.stsci.edu/hlsp/cdips}}.

@@

\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\subfloat{
			\includegraphics[width=0.49\textwidth]{f2a.pdf}
			\includegraphics[width=0.49\textwidth]{f2b.pdf}
		}
	\end{center}
	\vspace{-0.7cm}
  \caption{ {\bf HR diagrams of NGC\,2516.}
    {\it Left}: Photometric data from Gaia EDR3; no reddening
    correction has been applied.  The core (black) shows a
    main-sequence and turnoff consistent with stars of a fixed age and
    metallicity.  The halo (blue) is similar, though the binary
    sequence is less defined.  The faintest M dwarfs in the core and
    halo are brighter than field stars (gray) with the same color,
    consistent with the cluster M dwarfs not having yet reached the
    main-sequence.  {\it Right}: PARSEC isochrones ([Fe/H]=+0.06) are
    overplotted at intervals of $\log_{10}(t/\mathrm{yr})=\{8, 8.25,
    8.5\}$.  The arrow represents the average reddening correction
    applied to the models.  The opacity of the M-dwarf model
    atmospheres is less well-calibrated for $M_\star \lesssim
    0.45M_\odot$ ($\approx$M2V; $Bp-Rp\approx$2.2); these model values
    are shown with dashed lines.
    \label{fig:hr}
  }
\end{figure*}

@@

        \vspace{-0.5cm}
		\subfloat{
			\includegraphics[width=0.49\textwidth]{f3c.pdf}
			\includegraphics[width=0.49\textwidth]{f3d.pdf}
		}
	\end{center}
	\vspace{-0.7cm}
  \caption{ {\bf TESS rotations periods and dereddened Gaia colors for \cn.}
    {\it Top row}: \cn\ split into its core and halo.  Sets
    $\mathcal{A}$ ({\it left columns}) and $\mathcal{B}$ ({\it right
    columns}) undergo successive degrees of automated cleaning,
    described in Section~\ref{subsubsec:cluster}.  The Pleiades
    \citep[120\,Myr;][]{rebull_rotation_2016a} and Praesepe
    \citep[650\,Myr;][]{douglas_poking_2017} are shown for reference.
    Stars in the core and the halo of NGC\,2516 appear
    gyrochronologically coeval with the Pleiades at $(Bp-Rp)_0$$<$1.2.
    There is a narrow range of color, from $1.2<(Bp-Rp)_0<1.7$, in
    which some NGC\,2516 members appear slightly elevated with respect
    to the Pleiades, suggesting an older age.  
    Stars redder than $(Bp-Rp)_0\gtrsim 2.2$ (later than M2V) 
    are fainter than the magnitude limit of our survey.
    {\it Bottom row}: \cn\
    colored by binarity indicators.  Sources with $\mathrm{RUWE}>1.2$
    appear in red (astrometric binaries; 11\% of stars).  Overplotted
    in orange are sources at least 0.3\,mag above an empirically fit
    cluster isochrone (photometric binaries; 20\% of stars).
		\label{fig:rot}
	}
\end{figure*}

We began our rotation analysis by considering all \nkinematic\
candidate members of \cn.  For each source, we retrieved all available
CDIPS light curves, on a per-sector basis.  This yielded \ncdips\
stars with at least one sector from a CDIPS light curve.  Each of
these stars is brighter than the CDIPS magnitude limit of
$Rp=16$\,mag.  \nkinematicrpltsixteen\ of the stars from the \cn\
source list had $Rp<16$.  The difference is due to 34 stars unique to
\citet{meingast_2021} which were not available at the time of the
CDIPS reductions, and a few stars falling on the TESS chip gaps.  The
$Rp=16$\,mag cutoff imposed during the CDIPS processing corresponds
roughly to $(Bp-Rp)_0$ of 2.2, or a spectral type of $\approx$M2V, at
the distance of \cn.

We removed systematic trends shared between all light curves on each
CCD in each sector individually, and stitched together the resulting
light curves before searching for rotation-induced periodicity.
Details regarding our detrending approach are presented in
Appendix~\ref{app:detrending}.  After applying an initial detrending
step aimed at removing systematic trends, we proceeding with a few
small cleaning steps to improve the purity of the rotation period
measurements: we masked 0.7 days at the beginning and end of each
spacecraft orbit, and ran a sliding standard-deviation rejection
window over the light curve, which removed any outlying points within
$\pm3\times$MAD of the median in each window.

We then measured the rotation period from the resulting light curve
using the periodogram implementations in \texttt{astrobase}, in
particular the Stellingwerf phase-dispersion minimization periodogram,
along with the more traditional Lomb-Scargle
\citep{lomb_1976,stellingwerf_period_1978,scargle_studies_1982,stellingwerf_period_2011,bhatti_astrobase_2018}.
We used the CDIPS aperture radius that, based on theoretical
expectations, was expected to give the optimal balance between light
from the target and background-light \citep{Sullivan_2015}.  This
typically resulted in an aperture radius of either 1 or 1.5 pixels.
We recorded the top five periodogram peaks from each method, and their
corresponding powers.  Finally, as a check on crowding, we recorded
the number of stars within the aperture with equal or greater
brightness than the target star, and with brightness within 1.25 and
2.5 TESS magnitudes of the target star.

To clean the resulting rotation period measurements, we devised a set
of automated cleaning criteria, which yielded sets we denote
$\mathcal{A}$ and $\mathcal{B}$.  For both sets, we considered
light curves for which the peak Lomb-Scargle periodogram period was
below 15 days, the normalized Lomb-Scargle power exceeded 0.08, and
for which no equal-brightness or greater companions were within the
aperture.  Beyond requiring that the target star be the brightest star
in the aperture, we also required that at most one companion with
brightness exceeding one tenth of the target star's brightness be
present in the aperture according to the Gaia DR2 source catalog.  For
set $\mathcal{A}$, this yielded \nautorotnumerator\ light curves, out
of \nautorotdenominator\ stars for which rotation periods might
plausibly have been detected.  For set $\mathcal{B}$, we additionally
required that the Lomb-Scargle and Stellingwerf phase-dispersion
methods yielded identical periods to within 10\%.  This yielded
\nautorotnumeratormatching\ light curves.

The top row of Figure~\ref{fig:rot} shows the resulting rotation
periods, with set $\mathcal{A}$ shown the left and $\mathcal{B}$ on
the right.  While set $\mathcal{A}$ is more complete than set
$\mathcal{B}$, this completeness comes at the expense of greater
contamination.  For instance, stars above the slow sequence (possible
field contaminants) for the most part disappear in set $\mathcal{B}$.
The ``void'' or ``gap'' underneath the slow sequence is also more
pronounced in set $\mathcal{B}$.  To facilitate a comparison against
the Pleiades and Praesepe, we used the rotation periods and dereddened
colors from Table~5 of \citet{curtis_rup147_2020}, which drew on data
from \citet{rebull_rotation_2016a} and \citet{douglas_k2_2019}
respectively.  The first order conclusion is that the Pleiades and
\cn\ appear gyrochronologically coeval for colors spanning
$0.5<(Bp-Rp)_0<1.2$.  The implications of this comparison for the age
of \cn\ are assessed in greater detail in Section~\ref{disc:absage}.

The bottom row of Figure~\ref{fig:rot} shows the same data, with the
points colored with indicators of binarity.  To assess astrometric
binarity, we used the renormalized unit weight error (RUWE) from Gaia
EDR3.  For plotting purposes, we labeled anything with RUWE exceeding
1.2 as an astrometric binary (11\% of the overall cluster sample; see
{\it e.g.}, \citealt{belokurov_unresolved_2020}).  To assess
photometric binarity, we fitted a spline to Figure~\ref{fig:hr} by
hand, and labelled any points brighter than 0.3\,mag above the cluster
sequence as photometric binaries.  This included 20\% of the overall
cluster sample.  Many of the resulting candidate binaries, though not
all of them, appear underneath the ``slow sequence''.  We speculate on
possible explanations in Section~\ref{disc:rotn}.

@@

\begin{figure*}[t]
	\begin{center}
		\subfloat{
			\includegraphics[width=0.5\textwidth]{f8a.pdf}
			\includegraphics[width=0.5\textwidth]{f8b.pdf}
		}
	\end{center}
	\vspace{-0.7cm}
  \caption{ 
  {\bf Members of \cn\ exist hundreds of parsecs from the core,
  and up to $\approx$5\kms\ from the core in tangential velocity.}
  Halo members that meet both kinematic and rotational membership
  criteria (Subset~$\mathcal{A}$) are shown in blue.  Halo members for which rotation periods
  should have been detectable, but were not detected, are shown in
  orange.  
  The plotted stars are restricted to $0.5<(Bp-Rp)_0<1.2$.
  The {\it left panels} show galactic coordinates and
  2D-tangential velocity space, after
  correcting for projection effects
  (Appendix~\ref{app:vproj}).
  The histogram on the right bins versus 3D separation from the core
  and 2D velocity difference.  Bins have width 25\,pc in position, and
  1.0\,km\,s$^{-1}$ in velocity.  {\bf Uncertainties in the fraction
  are taken as the quadrature sum of the rel unc of the numerator and
  denominator (FIXME)}
  \label{fig:physical_x_rotn}
	}
\end{figure*}

@@

How far away from the core, in position and velocity space, does the
halo extend?  We can explore this by crossmatching the stars with
detected rotation periods against our original target list of
\nkinematic\ Gaia DR2 kinematic candidate members.  For a fair
comparison, we consider only the stars with $0.5<(Bp-Rp)_0<1.2$ for
which our TESS pipeline successfully made detrended light curves.  In
other words---the base sample is the stars for which we could have
plausibly detected a rotation period.  Faint stars with $Rp>16$ were
beyond our light curve selection limit, and crowded stars ({\it e.g.,}
near saturated stars in the cluster center) are similarly omitted.  We
selected the color limit above based on Figure~\ref{fig:rot}, since
the density of period detections seems to decrease once $(Bp-Rp)_0
\gtrsim 1.25$, {\it i.e.}, for spectral types later than $\approx$K4V.
Our expectation is that completion effects become important for
fainter stars.

@@

Gaia-ESO selected candidate \cn\ members to observe with the GIRAFFE
and UVES spectrographs based on previously reported literature members
and publicly available photometry.  Since the existence of the \cn\
halo was not known at the time of target selection, very few ``halo''
stars are in the sample.  For \cn, \citet{randich_gaiaeso_2018}
ultimately reported stellar parameters (including lithium equivalent
widths and metallicities) for 796 stars that they considered possible
\cn\ members.  Cross-matching against our kinematic list of
\nkinematic\ candidate members by position and imposing a 0$\farcs$5
maximum separation limit yields 492 kinematic members for which a
Gaia-ESO spectrum was acquired and stellar parameters were reported.
15 of these (all with $(Bp-Rp)_0 > 2.0$) are spurious matches based on
the Gaia color and GES effective temperature, and we remove them.
This yields 477 stars, of which 436 are in the core, and 41 are in the
halo.  The lop-sided ratio is due to the Gaia-ESO selection function.
We verified by directly querying the ESO archive that
\citet{randich_gaiaeso_2018} indeed included $>99\%$ of the available
Gaia-ESO spectra as part of their \cn\ sample, and so we opted to use
their results rather than repeating their analysis.

The GALAH DR3 target selection is discussed in depth by
\citet{buder_galah_2020}.  The relevant aspects for our analysis are
that the ``main'' survey targeted $12<V<14$ stars at
$\delta<+10^\circ$, provided the stars were at least ten degrees from
the galactic plane.  The survey is spatially
inhomogeneous\footnote{See the footprint at
\url{https://www.galah-survey.org/news/announcing-galah-dr3}}. Special
targeting for stars in the TESS southern continuous viewing zone, and
for known open cluster members was also performed.  We identified the
candidate \cn\ members for which spectra had been obtained by
searching the \texttt{GALAH\_DR3\_main\_allstar\_v1} catalog, after
excluding stars with the stellar parameter bit flags 1, 2, 3.  This
excludes spectra with unreliable broadening, low S/N, and unreliable
wavelength solutions (see Table~6 of \citealt{buder_galah_2020}).
Since our main focus is measuring equivalent widths for the
Li\,6708\,\AA doublet, these were the most relevant bit flags.  Of our
\nkinematic\ candidate \cn\ members, 107 had spectra in GALAH DR3.  51
were in the ``core''; 56 were in the ``halo''.  78 had ``finite''
lithium, i.e., no lithium flag set.

@@

\subsection{Rotation at 100-200\,Myr}
\label{disc:rotn}

From their rotation period analysis of \cn,
\citet{fritzewski_rotation_2020} have recently solidified our
understanding that at ages between 100 and 200\,Myr, young open
clusters ({Blanco\,1, the Pleiades, Psc\,Eri, M\,35, and M\,50) all
seem to show similar rotation period distributions.  The FGK stars
show a ``slow sequence'', a population of ``rapid rotators'', and
in some cases stars in the ``void'' between the two
populations ({\it e.g.,}, Psc\,Eri and \cn\ show such populations).
The late M dwarfs with detected rotation periods tend to rotate very
rapidly ($<3$\,days) --- these are unfortunately beyond the magnitude
limit of the present analysis.  In this section, we briefly discuss
these features, and focus on the main difference between \cn\
and the Pleiades: the longer maximum rotation periods seen in the
K-dwarfs.

\subsubsection{The slow and fast sequences}
The existence of a slow sequence is the main premise of gyrochronology
\citep[{\it e.g.},][]{barnes_rotational_2003}.  While the initial
spindown relation ($P_{\rm rot}\sim t^{1/2}$) proposed by
\citet{skumanich_time_1972} has not worked in detail, the fact that it
is simple and can be understood to first order through magnetic
breaking \citep{weber_angular_1967} has led to more systematic efforts
at age-dating stars through their rotation periods.

The other features of the rotation-color diagram are in some sense
affronts to the magnetic breaking model, and the general question of
what produces the rapid rotators has not yet been definitively
answered.  Models that introduces variable core-envelope decoupling
timescales can reproduce each type of population
\citep{Irwin_NGC2516_2007}, but the question then shifts to why these
decoupling timescales should be differ for stars that are otherwise
identical.  Another possible explanation invokes a spontaneous and
random change of magnetic field topology that cause the dynamo to
couple more strongly with the stellar wind
\citep{brown_metastable_2014}.  Other processes could be external the
star entirely: \citet{qureshi_signature_2018} for instance have
reported that mergers between giant planets and the star could explain
a non-neglible fraction of the rapid rotators.
